---
title: "Generate plots"
author: "Sabrina Xie"
date: "2/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

DON'T KNIT!!!!

Run Probability.Sampling.BASE.Rmd chunks 1-6, then run chunk for plot(s) you want

List of indicators:
Indicator=case_when(
           Indicator=="phos_trophic"~"Phosphorus", 
           Indicator=="chla_trophic"~"Chlorophyll-a", 
           Indicator=="TP_threshold"~"Total phosphorus", 
           Indicator=="TN_threshold"~"Total nitrogen", 
           Indicator=="CHLA_threshold"~"Total chlorophyll", 
           Indicator=="microcystin"~"Microcystin",
           Indicator=="d.oxygen"~"Dissolved oxygen",
           Indicator=="N_LIMIT"~"Nutrient limitation", 
           Indicator=="secchi"~"Secchi", 
           Indicator=="zebra"~"Zebra mussel susceptibility", 
           Indicator=="conductance"~"Hardness", 
           Indicator=="color"~"Color", 
           Indicator=="ph"~"pH", 
           Indicator=="leech"~"Nutrient-color status", 
           Indicator=="chloride"~"Chloride")

# Trophic State Indicators

```{r trophic, fig.height=3, fig.width=6}
vars <- c("Phosphorus","Chlorophyll-a","Secchi")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
```{r nutrient limitation, fig.height=1, fig.width=2}
vars <- c("Nutrient limitation")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
```{r leech, fig.height=1, fig.width=2}
vars <- c("Nutrient-color status")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```


# Chemical 1

```{r ph, fig.height=1, fig.width=2}
vars <- c("pH")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	

```{r oxygen, fig.height=1, fig.width=2}
vars <- c("Dissolved oxygen")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
  
```{r nutrients, fig.height=2, fig.width=4}
vars <- c("Total nitrogen","Total phosphorus")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
	
# Chemical 2

```{r hardness, fig.height=1, fig.width=2}
vars <- c("Hardness")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
```{r alkalinity, fig.height=1, fig.width=2}
vars <- c("pH")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
```{r ca, fig.height=1, fig.width=2}
vars <- c("Zebra mussel susceptibility")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
```{r color, fig.height=1, fig.width=2}
vars <- c("Color")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	
```{r chloride, fig.height=1, fig.width=2}
vars <- c("Chloride")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```


# Human Use

```{r chla, fig.height=1, fig.width=2}
vars <- c("Total chlorophyll")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```
	

```{r microcystin, fig.height=1, fig.width=2}
vars <- c("Microcystin")

ggplot(table[table$Indicator %in% vars,],aes(y=Estimate.P,x=Category,fill=Category))+
  geom_col()+
  ylim(0,100)+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  coord_flip()+
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title=element_blank())+
  facet_wrap(.~Indicator,scales = "free_y",ncol=1,strip.position = "left")+
  labs(y="Percent of Total")

```

