---
title: "Probability.Results.BASE"
author: "Sabrina Xie"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y, %X')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Probability analysis

## Load packages

```{r packages}

library(tidyverse)
library(huxtable)
library(ggplot2)
library(lubridate)
library(spsurvey) #This code was written for spsurvey v5.0.0

```

## Load data

To run cat_analysis(), the dataframe will need to include columns for the categorical variable you are interested in, weight for each site, xcoordinate and ycoordinate. Optionally, you can also include a subpopulation or stratum column to separate results and site IDs for two-stage analysis (for example, if there are repeated sites).

In this script, weight is adjusted because of the proportional design. In the future, this can be excluded as the design should include weights already.

I am designating relevant variables from the dataset as well as creating new variables: mean dissolved oxygen from 0-2m, epilimentic means for ph and spconductance, total nitrogen and DIN:TP ratio.

```{r data prep}

# retrieve raw data from database
setwd("~/OneDrive - New York State Office of Information Technology Services/Rscripts/Current")
source("new_database/Reading.LMAS.Data.R")
setwd("~/OneDrive - New York State Office of Information Technology Services/Rscripts/Probability.Sampling/")

rm(list=setdiff(ls(), c('newdata')))

#list relevant variables
vars_to_keep <- c('CHLOROPHYLL A_OW_TOTAL',
                  'PHOSPHORUS, TOTAL_OW_TOTAL',
                  "MICROCYSTIN_OW_NA",
                  "NITROGEN, NITRATE (AS N)_OW_TOTAL",
                  "NITROGEN, NITRATE-NITRITE_OW_TOTAL",
                  "NITROGEN, KJELDAHL, TOTAL_OW_TOTAL",
                  "NITROGEN, TOTAL_OW_TOTAL",
                  "DISSOLVED OXYGEN_DP_NA",
                  "TRUE COLOR_OW_TOTAL",
                  "PH_DP_NA",
                  "SPECIFIC CONDUCTANCE_DP_NA",
                  "TEMPERATURE_DP_NA",
                  "DEPTH, SECCHI DISK DEPTH_SD_NA",
                  "CALCIUM_OW_TOTAL",
                  "CHLORIDE_OW_TOTAL")

lab<-newdata %>%
  filter(SAMPLE_DATE>'2020-01-01') %>%
  mutate(combined=paste(CHARACTERISTIC_NAME,
                        INFORMATION_TYPE,
                        RSLT_RESULT_SAMPLE_FRACTION,
                        sep = "_"))  %>%
  select(LAKE_HISTORY_ID,
         SAMPLE_DATE,
         combined,
         RSLT_RESULT_VALUE,
         RSLT_LABORATORY_QUALIFIER,
         RSLT_VALIDATOR_QUALIFIER,
         RSLT_PROFILE_DEPTH) %>%
  mutate(RSLT_RESULT_VALUE=ifelse(!is.na(RSLT_LABORATORY_QUALIFIER)&(RSLT_LABORATORY_QUALIFIER=="U"|RSLT_LABORATORY_QUALIFIER=="UE"),"0",RSLT_RESULT_VALUE),
         RSLT_RESULT_VALUE=as.numeric(RSLT_RESULT_VALUE)) %>%
  filter(!is.na(RSLT_RESULT_VALUE),
         is.na(RSLT_VALIDATOR_QUALIFIER)|(RSLT_VALIDATOR_QUALIFIER!="R"),
         combined %in% vars_to_keep) %>%
  select(LAKE_HISTORY_ID,SAMPLE_DATE,combined,RSLT_RESULT_VALUE,RSLT_PROFILE_DEPTH) %>%
  distinct(LAKE_HISTORY_ID,SAMPLE_DATE,combined,RSLT_PROFILE_DEPTH,.keep_all = TRUE) %>%
  rename(LAKE_ID=LAKE_HISTORY_ID,
         chemical_name=combined,
         result_value=RSLT_RESULT_VALUE,
         profile_depth=RSLT_PROFILE_DEPTH)

# dissolved oxygen:
# keep DO values between 0 and 2m depth
lab2 <- lab %>% 
  mutate(keep=case_when(
    !(chemical_name=="DISSOLVED OXYGEN_DP_NA") ~ "no",
    chemical_name=="DISSOLVED OXYGEN_DP_NA" & profile_depth <= 2 ~ "yes",
    chemical_name=="DISSOLVED OXYGEN_DP_NA" & profile_depth > 2 ~ "no")) %>%
  filter(keep!="no") %>%
  select(-keep)

# mean DO for each lake each date
DO <- aggregate(lab2$result_value,list(lab2$LAKE_ID,lab2$SAMPLE_DATE),FUN=mean) %>% 
  mutate(chemical_name="DISSOLVED OXYGEN_epi") %>% 
  rename(LAKE_ID=Group.1,
         SAMPLE_DATE=Group.2,
         result_value=x)

#thermocline
thermocline<-lab %>% 
  filter(chemical_name=='TEMPERATURE_DP_NA') %>% 
  mutate(thermocline=NA)
LID<-thermocline$LAKE_ID[1]
date<-thermocline$SAMPLE_DATE[1]
depth<-thermocline$profile_depth[1]
temp<-thermocline$result_value[1]

for(i in seq(nrow(thermocline))){
  current<-thermocline[i,]
  if(current$LAKE_ID==LID&current$SAMPLE_DATE==date){
      depth=current$profile_depth
      if((temp-current$result_value)>1){
        thermocline$thermocline[i]<-current$profile_depth
      }
      temp=current$result_value
  }
  else{
    LID=current$LAKE_ID
    date=current$SAMPLE_DATE
    depth=current$profile_depth
    temp=current$result_value
  }
}

thermocline<-thermocline %>%  #pull the lowest value for all
  group_by(LAKE_ID,SAMPLE_DATE) %>% 
  mutate(thermocline=min(thermocline, na.rm=T)) %>% 
  ungroup() %>% 
  mutate(thermocline=ifelse(thermocline==Inf,NA,thermocline)) %>% 
  select(LAKE_ID,SAMPLE_DATE,thermocline,profile_depth) %>% 
  filter(!is.na(thermocline)) %>%   #remove those without a thermocline
  distinct()

#epilimentic means for ph and spconductance
epi<-lab %>% 
  inner_join(thermocline,by=c('LAKE_ID','SAMPLE_DATE','profile_depth')) %>% 
  filter(chemical_name %in% c("PH_DP_NA","SPECIFIC CONDUCTANCE_DP_NA")) %>% 
  mutate(result_value = ifelse(chemical_name=="PH_DP_NA",(10^result_value),result_value)) %>% 
  distinct() %>% 
  arrange(LAKE_ID,SAMPLE_DATE,chemical_name,profile_depth) 

epi<-epi %>% 
  filter(profile_depth<=thermocline) %>% 
  distinct() %>% 
  group_by(LAKE_ID,SAMPLE_DATE,chemical_name) %>% 
  summarize(Mean=mean(result_value,na.rm=TRUE),
            n=n()) %>% 
  filter(n>2) %>% 
  select(LAKE_ID,SAMPLE_DATE,chemical_name,Mean)

epi<-epi %>% 
  mutate(Mean=ifelse(chemical_name=="PH_DP_NA",log10(Mean),Mean)) %>% 
  rename(result_value=Mean) %>% 
  mutate(chemical_name=case_when(
    chemical_name=="PH_DP_NA"~"PH_epi",
    chemical_name=="SPECIFIC CONDUCTANCE_DP_NA"~"SPECIFIC CONDUCTANCE_epi"
  ))

# read in site data
sites<-read.csv("Probability_Based_Sites_2020_2021.csv")
sites<-sites %>% 
  rename(siteID=SITE_ID,
         xcoord=LON_DD83,
         ycoord=LAT_DD83,
         Eval_Status=EvalStatus) %>% 
  filter(Eval_Status!="") %>%    #removing sites that we haven't yet evaluated
  mutate(Eval_Status=trimws(Eval_Status))

# restrict to only the data in the probability study and spread the data
att<-merge(lab,epi,by=c("LAKE_ID","SAMPLE_DATE","chemical_name"),all=TRUE) %>% 
  mutate(result_value=ifelse(is.na(result_value.x),result_value.y,result_value.x)) %>% 
  select(-result_value.x,-result_value.y)

att<-merge(att,DO,by=c("LAKE_ID","SAMPLE_DATE","chemical_name"),all=TRUE) %>% 
  mutate(result_value=ifelse(is.na(result_value.x),result_value.y,result_value.x)) %>% 
  select(-result_value.x,-result_value.y) 
  
att<-merge(att,sites,by=c('LAKE_ID'),all.y = TRUE) %>% distinct()
att<-att %>% spread(chemical_name,result_value,fill = NA)

# Create a Target/NonTarget status variable
att<-att %>% 
  mutate(statusTNT="Target",
         statusTNT=ifelse(Eval_Status=="NonTarget","NonTarget",statusTNT))

# get duplicates (multiple samples of same lake)
n_occur <- data.frame(table(att$FID))
dupes <- att[att$FID %in% n_occur$Var1[n_occur$Freq > 1],]

# keep most recent august sample
m_att <- att %>%
  mutate(keep=case_when(
    FID %in% dupes$FID & month(SAMPLE_DATE) != 9 & !is.na(`CHLOROPHYLL A_OW_TOTAL`) ~ "yes",
    !(FID %in% dupes$FID) ~ "yes")) %>%
  arrange(FID,desc(SAMPLE_DATE)) %>%
  filter(keep=="yes")

m_occur <- data.frame(table(m_att$FID))
m_att <- m_att[!duplicated(m_att$FID),]

att <- m_att %>%
  distinct()

# add adjusted weight variable
att <- att %>% 
  mutate(WgtAdj=case_when(
         PROB_CAT=="(1,4]"~(1828/29),
         PROB_CAT=="(4,10]"~(2490/15),
         PROB_CAT=="(10,20]"~(1003/18),
         PROB_CAT=="(20,50]"~(616/20),
         PROB_CAT==">50"~(500/33)))

att <- att %>% 
  # create total nitrogen
  mutate(`NITROGEN, TOTAL`=case_when(
    !is.na(`NITROGEN, KJELDAHL, TOTAL_OW_TOTAL`) ~
      (`NITROGEN, NITRATE-NITRITE_OW_TOTAL`+`NITROGEN, KJELDAHL, TOTAL_OW_TOTAL`),
    is.na(`NITROGEN, KJELDAHL, TOTAL_OW_TOTAL`) ~ `NITROGEN, TOTAL_OW_TOTAL`)) %>%
  # DIN:TP
  mutate(DINTP=`NITROGEN, NITRATE (AS N)_OW_TOTAL`/`PHOSPHORUS, TOTAL_OW_TOTAL`)

```

## Creating thresholds

Create any desired thresholds. I use a case_when() statement.

```{r thresholds}

att<-att %>% 
  # remove fields in the sites table that aren't relevant
  select(-Accessible,-Comments,-Contact,-STATUS) %>%
  # trophic status
  mutate(phos_trophic=case_when(
    `PHOSPHORUS, TOTAL_OW_TOTAL`<=0.01 ~ "Oligo",
    between(`PHOSPHORUS, TOTAL_OW_TOTAL`,0.01,0.02) ~ "Meso",
    `PHOSPHORUS, TOTAL_OW_TOTAL`>=0.02 ~ "Eu")) %>%
  mutate(chla_trophic=case_when(
    `CHLOROPHYLL A_OW_TOTAL`<=2 ~ "Oligo",
    between(`CHLOROPHYLL A_OW_TOTAL`,2,8) ~ "Meso",
    `CHLOROPHYLL A_OW_TOTAL`>=8 ~ "Eu")) %>%
  # EPA thresholds
  mutate(TP_threshold=case_when(
    `PHOSPHORUS, TOTAL_OW_TOTAL`<=0.016 ~ "Good",
    between(`PHOSPHORUS, TOTAL_OW_TOTAL`, 0.016, 0.0279) ~ "Fair",
    `PHOSPHORUS, TOTAL_OW_TOTAL`>=0.0279 ~ "Poor")) %>%
  mutate(TN_threshold=case_when(
    `NITROGEN, TOTAL`<=0.428 ~ "Good",
    between(`NITROGEN, TOTAL`, 0.428, 0.655) ~ "Fair",
    `NITROGEN, TOTAL`>=0.655 ~ "Poor")) %>%
  mutate(CHLA_threshold=case_when(
    `CHLOROPHYLL A_OW_TOTAL`<=4.52 ~ "Good",
    between(`CHLOROPHYLL A_OW_TOTAL`, 4.52, 8.43) ~ "Fair",
    `CHLOROPHYLL A_OW_TOTAL`>=8.43 ~ "Poor")) %>%
  # microcystin
  mutate(microcystin=case_when(
    is.na(`MICROCYSTIN_OW_NA`) ~"Non-detect",
    `MICROCYSTIN_OW_NA`<8 ~ "Microcystin detected",
    `MICROCYSTIN_OW_NA`>=8 ~ "Most disturbed")) %>%
  # dissolved oxygen
  mutate(d.oxygen=case_when(
    `DISSOLVED OXYGEN_epi`<=3 ~ "Poor",
    between(`DISSOLVED OXYGEN_epi`, 3, 5) ~ "Fair",
    `DISSOLVED OXYGEN_epi`>=5 ~ "Good")) %>% 
  # nutrient limitation
  mutate(N_LIMIT=case_when(
    `NITROGEN, TOTAL`/`PHOSPHORUS, TOTAL_OW_TOTAL`<=10 ~ "N-limited",
    between(`NITROGEN, TOTAL`/`PHOSPHORUS, TOTAL_OW_TOTAL`, 10, 20) ~ "Co-limited",
    `NITROGEN, TOTAL`/`PHOSPHORUS, TOTAL_OW_TOTAL`>=20 ~ "P-limited")) %>% 
  # secchi
  mutate(secchi=case_when(
    `DEPTH, SECCHI DISK DEPTH_SD_NA`<=2 ~ "Eu",
    between(`DEPTH, SECCHI DISK DEPTH_SD_NA`, 2, 5) ~ "Meso",
    `DEPTH, SECCHI DISK DEPTH_SD_NA`>=5 ~ "Oligo")) %>% 
  # zebra mussels
  mutate(zebra=case_when(
    CALCIUM_OW_TOTAL<=10 ~ "Not susceptible",
    between(CALCIUM_OW_TOTAL, 10, 20) ~ "May be susceptible",
    CALCIUM_OW_TOTAL>=20 ~ "Highly susceptible")) %>% 
  # cslap (conductance, color, epilimnetic ph)
  mutate(conductance=case_when(
    `SPECIFIC CONDUCTANCE_epi`<=125 ~ "Soft",
    between(`SPECIFIC CONDUCTANCE_epi`, 125, 250) ~ "Average",
    `SPECIFIC CONDUCTANCE_epi`>=250 ~ "Hard")) %>%
  mutate(color=case_when(
    `TRUE COLOR_OW_TOTAL`<=10 ~ "Uncolored",
    between(`TRUE COLOR_OW_TOTAL`, 10, 30) ~ "Weak",
    `TRUE COLOR_OW_TOTAL`>=30 ~ "High")) %>% 
  mutate(ph=case_when(
    `PH_epi`<=6.5 ~ "Acidic",
    between(`PH_epi`, 6.5, 7.5) ~ "~Neutral",
    between(`PH_epi`, 7.5, 8.5) ~ "Slightly alk",
    `PH_epi`>=8.5 ~ "Highly alk")) %>% 
  #leech
  mutate(leech=case_when(
    `PHOSPHORUS, TOTAL_OW_TOTAL`<=0.030 & `TRUE COLOR_OW_TOTAL`<=20 ~ "Blue",
    `PHOSPHORUS, TOTAL_OW_TOTAL`>0.030 & `TRUE COLOR_OW_TOTAL`<=20 ~ "Green",
    `PHOSPHORUS, TOTAL_OW_TOTAL`<=0.030 & `TRUE COLOR_OW_TOTAL`>20 ~ "Brown",
    `PHOSPHORUS, TOTAL_OW_TOTAL`>0.030 & `TRUE COLOR_OW_TOTAL`>20 ~ "Murky"
  )) %>% 
  #chloride
  mutate(chloride=case_when(
    `CHLORIDE_OW_TOTAL` <= 25 ~ "Good",
    between(`CHLORIDE_OW_TOTAL`,25,100) ~ "Fair",
    `CHLORIDE_OW_TOTAL` >= 100 ~ "Poor",
    Eval_Status=="Target_Sampled" & is.na(`CHLORIDE_OW_TOTAL`) ~ "Good"))

rm(list=setdiff(ls(), c('newdata',"att")))

```

## Analysis

Here I perform the analysis using cat_analysis(). The relevant categorical variable / threshold should be input as the "vars" argument. "vars" can be a list of all categories you are interested in, which would create a very big dataframe.

In the future, the "weight" argument should be part of the site design and might not be called "WgtAdj".

```{r analysis}

#list variables you are interested in and defined above
vars <- c("phos_trophic", "chla_trophic", "TP_threshold", "TN_threshold", "CHLA_threshold", "microcystin", "d.oxygen", "N_LIMIT", "secchi", "zebra", "conductance", "color", "ph", "leech", "chloride")

#analysis
CatExtent <- cat_analysis(
  dframe=att,
  vars=vars, 
  subpops = , #for example could separate by area category or year
  siteID = "siteID",
  weight = "WgtAdj", #name will probably change in future
  xcoord = "xcoord",
  ycoord = "ycoord")

table <- CatExtent %>% 
  select(Indicator,
         Category,
         Estimate.P,
         LCB95Pct.P,
         UCB95Pct.P) %>% 
  filter(Category!="Total") %>% 
  mutate(Category=factor(Category, levels=c("Good","Fair","Poor",
                                            "Blue","Green","Brown","Murky",
                                            "Uncolored","Weak","High",
                                            "Acidic","~Neutral","Slightly alk","Highly alk",
                                            "Soft","Average","Hard",
                                            "Not susceptible","May be susceptible","Highly susceptible",
                                            "Oligo","Meso","Eu",
                                            "N-limited","Co-limited","P-limited",
                                            "Non-detect","Microcystin detected","Most disturbed")))

hux <- as_hux(table) #I use huxtable because kable doesn't work on my computer but that is also fine 
number_format(hux) <- 2
theme_plain(hux)

```

## Plots

Here I have plotted all the results together, but you could plot them individually.

```{r plots, fig.height=10, fig.width=10}

ggplot(table,aes(x=Category,y=Estimate.P)) +
  geom_point()+
  geom_errorbar(aes(ymin=LCB95Pct.P,ymax=UCB95Pct.P),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(.~Indicator,scales = "free")+
  ylim(0,100)+
  labs(title="Condition estimates across NYS Ponded Waters",y="Percent of Total",x="Condition category")

```

## Other plots

To compare to NLA data, run the same analysis code above for NLA relevant data. Create a "study" column and plot grouped/colored by "study".

To create maps: 

```{r map}

#read in lat/long for lakes
locs <- att

#fetch map
library(ggmap)
library(maps)

states <- map_data("state")
ny <- subset(states, region %in% c("new york"))

locs$CHLA_threshold <- factor(locs$CHLA_threshold, levels=c("Good","Fair","Poor"))

print(status <- ggplot() + 
  geom_polygon(data=ny,aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3) +
  geom_point(data=locs,aes(x=xcoord,y=ycoord,color=(Eval_Status=="Target_Sampled")))+
  guides(fill=FALSE))  # do this to leave off the color legend)

print(CHLA <- ggplot() + 
  geom_polygon(data=ny,aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3) +
  geom_point(data=locs,aes(x=xcoord,y=ycoord,color=CHLA_threshold))+
  guides(fill=FALSE))  # do this to leave off the color legend)

```