---
title: "Probability Sampling Results NLA 2017"
author: "Alene Onion and Tony Olsen of USEPA (modified by Sabrina Xie)"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y, %X')`"
output: html_document
---

```{r GlobalOptions}
options(knitr.duplicate.label = 'allow')
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# {.tabset}
Doing the probability sampling results analysis with NLA data of lakes with area >6.5 acres

## Prepping data

Merging relevant raw data files and filtering

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
################################################################################################################
#reading in and formatting raw data
################################################################################################################

library(tidyverse)
library(huxtable)
library(ggplot2)
library(lubridate)

# read in NLA data
setwd("~/OneDrive - New York State Office of Information Technology Services/Rscripts/Probability.Sampling/")

chem <- read.csv("nla_2017_water_chemistry_chla-data.csv")

chem <- chem %>% 
  filter(ANALYTE %in% c("NTL","CHLA","PTL"),
         NARS_FLAG == "") %>% 
  select(UID,SITE_ID,DATE_COL,ANALYTE,RESULT) %>% 
  pivot_wider(names_from= ANALYTE ,values_from = RESULT,) %>% 
  rename(TN=NTL,TP=PTL)

oxygen <- read.csv("nla_2017_profile-data.csv")

oxygen <- oxygen %>% 
  select(UID,SITE_ID,DATE_COL,DEPTH,OXYGEN) %>% 
  # dissolved oxygen:
  # keep DO values between 0 and 2m depth
  mutate(keep=case_when(
    DEPTH <= 2 ~ "yes",
    DEPTH > 2 ~ "no")) %>%
  filter(keep=="yes") %>%
  select(-keep)

# mean DO for each lake
DO <- aggregate(oxygen$OXYGEN,list(oxygen$SITE_ID,oxygen$DATE_COL),FUN=mean)
# add back in
oxygen <- merge(oxygen,DO,by.x=c("SITE_ID","DATE_COL"),by.y=c("Group.1","Group.2"),all.x=TRUE)
oxygen$x <- as.numeric(oxygen$x)
oxygen <- oxygen %>% 
  select(-c(DEPTH,OXYGEN)) %>%
  rename(OXYGEN=x) %>% 
  distinct()

sites <- read.csv("nla_2017_site_information-data.csv")

cats.full <- data.frame(
  AREA_CAT6=unique(sites$AREA_CAT6),
  n=as.double(1,2,3,4,5))

for(i in 1:5){
  b <- cats.full[i,1]
  a <- att %>%
    filter(AREA_CAT6==b)
  cats.full[i,2] <- nrow(a)
}

sites <- sites %>% 
  filter(AREA_HA>2.63) %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,LAT_DD83,LON_DD83) %>% 
  rename(xcoord=LAT_DD83,
         ycoord=LON_DD83)

att <- merge(chem,oxygen,by=c("UID","SITE_ID","DATE_COL"),all.y=TRUE,all.x=TRUE)
att <- merge(att,sites,by=c("UID","SITE_ID"))
att$CHLA <- as.numeric(att$CHLA)
att$TN <- as.numeric(att$TN)
att$TP <- as.numeric(att$TP)
att$OXYGEN <- as.numeric(att$OXYGEN)
  
# selecting parameters and adding trophic status

att<-att %>% 
  # EPA thresholds
  mutate(TP_threshold=case_when(
    TP<=0.016 ~ "Good",
    between(TP, 16, 27.9) ~ "Fair",
    TP>=0.0279 ~ "Poor")) %>%
  mutate(TN_threshold=case_when(
    TN<=0.428 ~ "Good",
    between(TN, 428, 655) ~ "Fair",
    TN>=0.655 ~ "Poor")) %>%
  mutate(CHLA_threshold=case_when(
    CHLA<=4.52 ~ "Good",
    between(CHLA, 4.52, 8.43) ~ "Fair",
    CHLA>=8.43 ~ "Poor")) %>%
  # dissolved oxygen
  mutate(d.oxygen=case_when(
    OXYGEN<=3 ~ "Poor",
    between(OXYGEN, 3, 5) ~ "Fair",
    OXYGEN>=5 ~ "Good"))

```

## Analysis

The start of the NYSAnalysis code from Tony Olsen

The results are weighted based on the distribution of the five size classes in the original data frame divided by the distribution in the samples evaluated. The purpose of this step is to account for oversampling in certain size classes.

PLEASE NOTE: the samples evaluated included the lakes that were actually sampled as well as the lakes which could not be sampled due to wetland status or access issues. After adjusting the weights, they should add up to the original frame size of 6437:

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

# Calculate the adjusted weights
# The numerator is the original frame size value
# the denominator is the actual number of samples evaluated in each category

#number of lakes in each probability category
cats <- data.frame(
  AREA_CAT6=unique(att$AREA_CAT6),
  n=as.double(1,2,3,4,5))

for(i in 1:5){
  b <- cats[i,1]
  a <- att %>%
    filter(AREA_CAT6==b)
  cats[i,2] <- nrow(a)
}

# add adjusted weight variable
att<-att %>% 
  mutate(WgtAdj=case_when(
         AREA_CAT6=="(1,4]"~(67/67),
         AREA_CAT6=="(4,10]"~(188/188),
         AREA_CAT6=="(10,20]"~(211/211),
         AREA_CAT6=="(20,50]"~(237/237),
         AREA_CAT6==">50"~(357/357)))

print(sum(att$WgtAdj))

framesize_LC <- c("EPA"=1060) 

```

## Comparison to NLA thresholds

The following code performs the analysis against EPA's condition thresholds as used in the National Lakes Assessments. It uses the number of sampled lakes in each condition category (for total phosphorus, total nitrogen and chlorophyll) to estimate the ratio of the 2583 lakes we could have sampled are in good, fair and poor condition.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

library(spsurvey)

# EPA nutrient threshold
#  Estimate assessment result
CatExtent <- cat_analysis(
  dframe=att,
  vars=c("TP_threshold","CHLA_threshold","TN_threshold"),
  subpops = ,
  siteID = "SITE_ID",
  weight = "WgtAdj",
  xcoord = "xcoord",
  ycoord = "ycoord")

#quickly calculate the percentages
totals<-CatExtent %>% 
  select(Indicator,Category,Estimate.U) %>% 
  rename(total=Estimate.U)
totals$total <- CatExtent$Estimate.U[4]
CatExtent<-merge(CatExtent,totals,by=c('Indicator','Category'),all.x = TRUE)

forplot<-CatExtent %>% 
  mutate(pct=(Estimate.U/total)*100,
         pct_lcb=(LCB95Pct.U/total)*100,
         pct_ucb=(UCB95Pct.U/total)*100) %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  filter(Category!="Total") %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="CHLA_threshold","ChlA",Indicator),
         Indicator=ifelse(Indicator=="TP_threshold","Phosphorus",Indicator),
         Indicator=ifelse(Indicator=="TN_threshold","Total Nitrogen",Indicator))

table <- forplot %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good", "Fair", "Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of NYS Ponded Waters across EPA thresholds",y="Percent of Total",x="EPA threshold")
```


Below the same analysis is performed using the number of sampled lakes in each microcystin condition category to estimate the ratio of the 2583 lakes we could have sampled that are least or most disturbed. All lakes were non-detects for microcystin. 

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
# Microcystin
#  Estimate assessment result
MicExtent <- cat_analysis(
  dframe=att,
  vars=c("microcystin"),
  subpops = ,
  siteID = "siteID",
  weight = "WgtAdj",
  xcoord = "xcoord",
  ycoord = "ycoord")

#quickly calculate the percentages
totals<-MicExtent %>% 
  select(Indicator,Category,Estimate.U) %>% 
  rename(total=Estimate.U)
MicExtent<-merge(MicExtent,totals,by=c('Indicator','Category'),all.x = TRUE)

forplot3<-MicExtent %>% 
  mutate(pct=(Estimate.U/total)*100,
         pct_lcb=(LCB95Pct.U/total)*100,
         pct_ucb=(UCB95Pct.U/total)*100) %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  filter(Category!="Total") %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="microcystin","Microsystin condition category",Indicator))

table <- forplot3 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Non-detect","Microcystin disturbed"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of microcystin conditions across NYS Ponded Waters",y="Percent of Total",x="Condition category")
```


Below, the analysis is done for each dissolved oxygen condition category.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
# Dissolved oxygen
#  Estimate assessment result
DOextent <- cat_analysis(
  dframe=att,
  vars=c("d.oxygen"),
  subpops = ,
  siteID = "siteID",
  weight = "WgtAdj",
  xcoord = "xcoord",
  ycoord = "ycoord")

#quickly calculate the percentages
totals<-DOextent %>% 
  select(Indicator,Category,Estimate.U) %>% 
  rename(total=Estimate.U)
totals$total <- DOextent$Estimate.U[4]
DOextent<-merge(DOextent,totals,by=c('Indicator','Category'),all.x = TRUE)

forplot4<-DOextent %>% 
  mutate(pct=(Estimate.U/total)*100,
         pct_lcb=(LCB95Pct.U/total)*100,
         pct_ucb=(UCB95Pct.U/total)*100) %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  filter(Category!="Total") %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="d.oxygen","Dissolved Oxygen",Indicator))

table <- forplot4 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good","Fair","Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of dissolved oxygen conditions across NYS Ponded Waters",y="Percent of Total",x="Condition category")
```

## Comparison to LMAS Samples

A plot below shows the distribution of the size classes included in the data frame. For comparison, I've plotted the proportion of NYS lakes sampled in each size class since 2010.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#Plotting data frame distribution as well as NYS sample distribution collected since 2010
#LMAS numbers retrieved in 2020

frame<-att %>% 
  select(PROB_CAT,AREA_CAT) %>% 
  rename(size_category=AREA_CAT) %>% 
  distinct() %>%
  mutate(Probability_Sites = case_when(
    endsWith(size_category, "(1,4]") ~ ((1828/6437)*100),
    endsWith(size_category, "(4,10]") ~ ((2490/6437)*100),
    endsWith(size_category, "(10,20]") ~ ((1003/6437)*100),
    endsWith(size_category, "(20,50]") ~ ((616/6437)*100),
    endsWith(size_category, ">50") ~ ((500/6437)*100))
    ) %>%
  mutate(LMAS_Sites = case_when(
    endsWith(size_category, "(1,4]") ~ (7.64),
    endsWith(size_category, "(4,10]") ~ (16.7),
    endsWith(size_category, "(10,20]") ~ (16.2),
    endsWith(size_category, "(20,50]") ~ (19.3),
    endsWith(size_category, ">50") ~ (40.1)
  )) %>%
  ungroup() %>%
  select(-PROB_CAT) %>% 
  gather(study,percentage,-size_category) %>% 
  arrange(study,size_category) %>% 
  mutate(percentage=as.numeric(percentage),
         size_category=factor(size_category,levels=c("(1,4]","(4,10]","(10,20]","(20,50]",">50")))

ggplot(frame, aes(x=size_category,y=percentage,fill=study)) +
  geom_bar(stat = "identity", position = position_dodge())+
  ylim(0,100)+
  labs(title="Size classes in the Probability Study versus LMAS samples since 2010",y="Percent of Total",x="Size Category (hectares)")
```

And the following plot shows the comparison of trophic distributions among LMAS most recent august samples since 2010 compared to the probability drawn samples.
NOTE: used ecoregion analysis to generate the LMAS numbers. These are the most recent August sample collected from each lake sampled since 2010 (a total of 473 unique lakes were sampled since 2010).

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

forplot5<-forplot2 %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="probability") %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="eutrophic",
          study="LMAS",
          percent=((215/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="mesotrophic",
          study="LMAS",
          percent=((182/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="oligotrophic",
          study="LMAS",
          percent=((76/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="eutrophic",
          study="LMAS",
          percent=((211/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="mesotrophic",
          study="LMAS",
          percent=((141/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="oligotrophic",
          study="LMAS",
          percent=((121/473)*100)) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

ggplot(forplot5, aes(x=Category,fill=study)) +
  geom_bar(aes(y=percent),stat = "identity", position = position_dodge())+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb), width=.05,
                 position=position_dodge(.9))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)
  labs(title="Trophic Classes in the Probability Study versus LMAS samples since 2010",y="Percent of Total",x="Trophic Classes")
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

# retrieve LMAS data in 2021 and count frequency of observations in each condition category
# lab2<-newdata %>%
#   filter(SAMPLE_DATE>'2010-01-01') %>%
#   mutate(combined=paste(CHARACTERISTIC_NAME,
#                         INFORMATION_TYPE,
#                         RSLT_RESULT_SAMPLE_FRACTION,
#                         sep = "_"))  %>%
#   select(LAKE_HISTORY_ID,
#          SAMPLE_DATE,
#          combined,
#          RSLT_RESULT_VALUE,
#          RSLT_LABORATORY_QUALIFIER,
#          RSLT_VALIDATOR_QUALIFIER,
#          RSLT_PROFILE_DEPTH) %>%
#   mutate(RSLT_RESULT_VALUE=ifelse(!is.na(RSLT_LABORATORY_QUALIFIER)&(RSLT_LABORATORY_QUALIFIER=="U"|RSLT_LABORATORY_QUALIFIER=="UE"),"0",RSLT_RESULT_VALUE),
#          RSLT_RESULT_VALUE=as.numeric(RSLT_RESULT_VALUE)) %>%
#   filter(!is.na(RSLT_RESULT_VALUE),
#          is.na(RSLT_VALIDATOR_QUALIFIER)|(RSLT_VALIDATOR_QUALIFIER!="R"),
#          combined %in% c('CHLOROPHYLL A_OW_TOTAL',
#                          'PHOSPHORUS, TOTAL_OW_TOTAL',
#                          "NITROGEN, NITRATE-NITRITE_OW_TOTAL",
#                          "NITROGEN, KJELDAHL, TOTAL_OW_TOTAL",
#                          "NITROGEN, TOTAL_OW_TOTAL",
#                          "DISSOLVED OXYGEN_DP_NA")) %>%
#   select(LAKE_HISTORY_ID,SAMPLE_DATE,combined,RSLT_RESULT_VALUE,RSLT_PROFILE_DEPTH) %>%
#   distinct(LAKE_HISTORY_ID,SAMPLE_DATE,combined,RSLT_PROFILE_DEPTH,.keep_all = TRUE) %>%
#   rename(LAKE_ID=LAKE_HISTORY_ID,
#          chemical_name=combined,
#          result_value=RSLT_RESULT_VALUE,
#          profile_depth=RSLT_PROFILE_DEPTH) %>%
#   # dissolved oxygen:
#   # keep DO values between 0 and 2m depth
#   mutate(keep=case_when(
#     !(chemical_name=="DISSOLVED OXYGEN_DP_NA") ~ "yes",
#     chemical_name=="DISSOLVED OXYGEN_DP_NA" & profile_depth <= 2 ~ "yes",
#     chemical_name=="DISSOLVED OXYGEN_DP_NA" & profile_depth > 2 ~ "no")) %>%
#   filter(keep!="no") %>%
#   select(-keep)
# 
# # mean DO for each lake
# DO <- aggregate(lab2$result_value,list(lab2$LAKE_ID),FUN=mean)
# # add back in
# lab2 <- merge(lab2,DO,by.x="LAKE_ID",by.y="Group.1",all.x=TRUE)
# lab2$x <- as.numeric(lab2$x)
# lab2 <- lab2 %>%
#   mutate(result=case_when(
#     chemical_name=="DISSOLVED OXYGEN_DP_NA" ~ x,
#     !(chemical_name=="DISSOLVED OXYGEN_DP_NA") ~ result_value)) %>%
#   select(-c(result_value,profile_depth,x)) %>%
#   rename(result_value=result) %>% 
#   distinct(LAKE_ID,SAMPLE_DATE,chemical_name,.keep_all = TRUE)
# 
# n_att<-lab2 %>% spread(chemical_name,result_value,fill = NA)
# n_att<-n_att %>%
#   # create total nitrogen
#   mutate(`NITROGEN, TOTAL`=case_when(
#     !is.na(`NITROGEN, KJELDAHL, TOTAL_OW_TOTAL`) ~
#       (`NITROGEN, NITRATE-NITRITE_OW_TOTAL`+`NITROGEN, KJELDAHL, TOTAL_OW_TOTAL`),
#     is.na(`NITROGEN, KJELDAHL, TOTAL_OW_TOTAL`) ~ `NITROGEN, TOTAL_OW_TOTAL`)) %>%
#   # trophic status
#   mutate(phos_trophic=case_when(
#     `PHOSPHORUS, TOTAL_OW_TOTAL`<=0.01 ~ "oligotrophic",
#     between(`PHOSPHORUS, TOTAL_OW_TOTAL`,0.01,0.02) ~ "mesotrophic",
#     `PHOSPHORUS, TOTAL_OW_TOTAL`>=0.02 ~ "eutrophic")) %>%
#   mutate(chla_trophic=case_when(
#     `CHLOROPHYLL A_OW_TOTAL`<=2 ~ "oligotrophic",
#     between(`CHLOROPHYLL A_OW_TOTAL`,2,8) ~ "mesotrophic",
#     `CHLOROPHYLL A_OW_TOTAL`>=8 ~ "eutrophic")) %>%
#   # EPA thresholds
#   mutate(TP_threshold=case_when(
#     `PHOSPHORUS, TOTAL_OW_TOTAL`<=0.016 ~ "Good",
#     between(`PHOSPHORUS, TOTAL_OW_TOTAL`, 0.016, 0.0279) ~ "Fair",
#     `PHOSPHORUS, TOTAL_OW_TOTAL`>=0.0279 ~ "Poor")) %>%
#   mutate(TN_threshold=case_when(
#     `NITROGEN, TOTAL`<=0.428 ~ "Good",
#     between(`NITROGEN, TOTAL`, 0.428, 0.655) ~ "Fair",
#     `NITROGEN, TOTAL`>=0.655 ~ "Poor")) %>%
#   mutate(CHLA_threshold=case_when(
#     `CHLOROPHYLL A_OW_TOTAL`<=4.52 ~ "Good",
#     between(`CHLOROPHYLL A_OW_TOTAL`, 4.52, 8.43) ~ "Fair",
#     `CHLOROPHYLL A_OW_TOTAL`>=8.43 ~ "Poor")) %>%
#   # dissolved oxygen
#   mutate(d.oxygen=case_when(
#     `DISSOLVED OXYGEN_DP_NA`<=3 ~ "Poor",
#     between(`DISSOLVED OXYGEN_DP_NA`, 3, 5) ~ "Fair",
#     `DISSOLVED OXYGEN_DP_NA`>=5 ~ "Good"))
# 
# 
# # get duplicates (multiple samples of same lake)
# n_occur <- data.frame(table(n_att$LAKE_ID))
# dupes <- n_att[n_att$LAKE_ID %in% n_occur$Var1[n_occur$Freq > 1],]
# 
# # # keep most recent august sample
# o_att <- n_att %>%
#   mutate(keep=case_when(
#     LAKE_ID %in% dupes$LAKE_ID & month(SAMPLE_DATE) != 9 & !is.na(`CHLOROPHYLL A_OW_TOTAL`) ~ "yes",
#     !(LAKE_ID %in% dupes$LAKE_ID) ~ "yes")) %>%
#   arrange(LAKE_ID,desc(SAMPLE_DATE)) %>%
#   filter(keep=="yes")
# 
# o_occur <- data.frame(table(o_att$LAKE_ID))
# o_att <- o_att[!duplicated(o_att$LAKE_ID),]
# 
# o_att <- o_att %>%
#   # remove extra columns
#   select(-c('CHLOROPHYLL A_OW_TOTAL',
#             'PHOSPHORUS, TOTAL_OW_TOTAL',
#             "NITROGEN, NITRATE-NITRITE_OW_TOTAL",
#             "NITROGEN, KJELDAHL, TOTAL_OW_TOTAL",
#             "NITROGEN, TOTAL_OW_TOTAL",
#             "NITROGEN, TOTAL"))
# 
# #number of lakes in each threshold category
# data.frame(table(o_att$TP_threshold))
# data.frame(table(o_att$TN_threshold))
# data.frame(table(o_att$CHLA_threshold))
# data.frame(table(o_att$d.oxygen))


forplot6<-forplot %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="probability") %>% 
  add_row(Indicator="ChlA",
          Category="Good",
          study="LMAS",
          percent=((306/752)*100)) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="LMAS",
          percent=((144/752)*100)) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="LMAS",
          percent=((302/752)*100)) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="LMAS",
          percent=((262/501)*100)) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="LMAS",
          percent=((82/501)*100)) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="LMAS",
          percent=((157/501)*100)) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="LMAS",
          percent=((186/441)*100)) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="LMAS",
          percent=((108/441)*100)) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="LMAS",
          percent=((147/441)*100)) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

forplot6$Category <- factor(forplot6$Category, levels=c("Good","Fair","Poor"))

ggplot(forplot6, aes(x=Category,fill=study)) +
  geom_bar(aes(y=percent),stat = "identity", position = position_dodge())+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb), width=.05,
                 position=position_dodge(.9))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="EPA condition categories in  Probability Study vs. LMAS samples since 2010",y="Percent of Total",x="EPA Condition Classes")
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
forplot7<-forplot4 %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="probability") %>% 
  add_row(Indicator="d.oxygen",
          Category="Good",
          study="LMAS",
          percent=((400/486)*100)) %>% 
  add_row(Indicator="d.oxygen",
          Category="Fair",
          study="LMAS",
          percent=((76/486)*100)) %>% 
  add_row(Indicator="d.oxygen",
          Category="Poor",
          study="LMAS",
          percent=((10/486)*100)) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

forplot7$Category <- factor(forplot7$Category, levels=c("Good","Fair","Poor"))

ggplot(forplot7, aes(x=Category,fill=study)) +
  geom_bar(aes(y=percent),stat = "identity", position = position_dodge())+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb), width=.05,
                 position=position_dodge(.9))+
  ylim(0,100)+
  labs(title="Dissolved oxygen condition in  Probability Study vs. LMAS samples since 2010",y="Percent of Total",x="D.O. Condition")
```


## Notes for future study:

To consider the influence of other strata (such as depth or ecoregion) keep the adjusted weights the same and use these categories to define sub populations. If we include two sub populations (see how large shallow lakes fair in trophic status) then it's sliced so thin that the confidence interval is really too large to be of any value. 

The question then is when is it valuable to stratify a study as opposed to using subpopulations in the post analysis? The answer is that it depends on the distribution of sub populations. If one of the sub populations is very small compared to the other, for example Long Island samples versus samples from other eco regions, then we would probably want to stratify the samples to ensure we get a representative set from Long Island. If the sub populations are relatively equal, then there's no reason to stratify.
