---
title: "Probability Sampling Results NH"
author: "Sabrina Xie"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y, %X')`"
output: html_document
---

```{r GlobalOptions}
options(knitr.duplicate.label = 'allow')
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# {.tabset}
Doing the probability sampling results analysis with NH lakes with area >6.5 acres

## Prepping data

Merging relevant raw data files and filtering

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
################################################################################################################
#reading in and formatting raw data
################################################################################################################

library(tidyverse)
library(huxtable)
library(ggplot2)
library(lubridate)

# read in NLA data

setwd("~/OneDrive - New York State Office of Information Technology Services/Rscripts/Probability.Sampling/")

att <- read.csv("New_Hampshire_Lakes_2017_Design_Status_20211027_KH_EPA.csv", na.strings=c(""," ","NA"))
  
# selecting parameters and adding trophic status

att<-att %>% 
  mutate(include=case_when(
    AREA_HA>2.63 ~ "yes",
    SITE_ID=TRUE ~ "no"))

framesize_LC <- c("NH"=3596) 

```

## Analysis

The start of the NYSAnalysis code from Tony Olsen

The results are weighted based on the distribution of the five size classes in the original data frame divided by the distribution in the samples evaluated. The purpose of this step is to account for oversampling in certain size classes.

## Comparison to NYS thresholds

The following code actually performs the analysis. It uses the number of sampled lakes in each trophic category (both for phosphorus and chlorophyll) to estimate the ratio of the 2583 lakes we could have sampled are eutrophic, mesotrophic, or eutrophic. 


PLEAE NOTE: this presumes that the lakes we didn't sample were random. A reasonable assumption but it's obviously preferable to sample all the lakes in the list to avoid this uncertainty. 

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

library(spsurvey)

# Size classes
SizeExtent <- cat_analysis(
  dframe=att,
  vars=c("AREA_CAT6"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXT",
  xcoord = "LON_DD_N83",
  ycoord = "LAT_DD_N83")
table <- SizeExtent %>% 
  filter(Subpopulation=="yes") %>%
  select(Category,nResp,Estimate.U,LCB95Pct.U,UCB95Pct.U)
hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

```

## Comparison to NLA thresholds

The following code performs the analysis against EPA's condition thresholds as used in the National Lakes Assessments. It uses the number of sampled lakes in each condition category (for total phosphorus, total nitrogen and chlorophyll) to estimate the ratio of all lakes in good, fair and poor condition.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

# EPA nutrient threshold
# Estimate assessment result
CatExtent <- cat_analysis(
  dframe=att,
  vars=c("EPA_Chla","EPA_TN","EPA_TP","DO"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXT",
  xcoord = "LON_DD_N83",
  ycoord = "LAT_DD_N83")

forplot<-CatExtent %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Subpopulation=="yes",Category!="Total") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="EPA_Chla","ChlA",Indicator),
         Indicator=ifelse(Indicator=="EPA_TP","Phosphorus",Indicator),
         Indicator=ifelse(Indicator=="EPA_TN","Total Nitrogen",Indicator),
         Indicator=ifelse(Indicator=="DO","Dissolved Oxygen",Indicator))

table <- forplot %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good", "Fair", "Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of Northern Appalachians ponded waters across EPA thresholds",y="Percent of Total",x="EPA threshold")
```


## Comparison to NY probability and LMAS

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

forplot6<-forplot %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NH") %>% 
  add_row(Indicator="ChlA", #NAP results
          Category="Good",
          study="NAP",
          percent=73.22,
          pct_lcb=64.45,
          pct_ucb=81.99) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="NAP",
          percent=7.96,
          pct_lcb=3.68,
          pct_ucb=12.23) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="NAP",
          percent=18.82,
          pct_lcb=11.26,
          pct_ucb=26.38) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="NAP",
          percent=64.20,
          pct_lcb=53.89,
          pct_ucb=74.51) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="NAP",
          percent=20.81,
          pct_lcb=12.76,
          pct_ucb=28.85) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="NAP",
          percent=15.00,
          pct_lcb=8.25,
          pct_ucb=21.74) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="NAP",
          percent=66.41,
          pct_lcb=55.32,
          pct_ucb=77.50) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="NAP",
          percent=25.15,
          pct_lcb=13.57,
          pct_ucb=36.73) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="NAP",
          percent=8.44,
          pct_lcb=4.23,
          pct_ucb=12.65) %>% 
  add_row(Indicator="ChlA", #nys prob results
          Category="Good",
          study="NY",
          percent=50.01,
          pct_lcb=33.05,
          pct_ucb=66.97) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="NY",
          percent=40.81,
          pct_lcb=16.57,
          pct_ucb=65.05) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="NY",
          percent=9.18,
          pct_lcb=2.88,
          pct_ucb=15.48) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="NY",
          percent=65.59,
          pct_lcb=42.85,
          pct_ucb=88.33) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="NY",
          percent=15.44,
          pct_lcb=5.85,
          pct_ucb=25.04) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="NY",
          percent=18.97,
          pct_lcb=6.22,
          pct_ucb=31.72) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="NY",
          percent=61.49,
          pct_lcb=37.27,
          pct_ucb=85.71) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="NY",
          percent=14.95,
          pct_lcb=4.34,
          pct_ucb=25.56) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="NY",
          percent=23.57,
          pct_lcb=6.99,
          pct_ucb=40.14)  %>% 
  add_row(Indicator="Dissolved Oxygen", #NLA
          Category="Good",
          study="NLA",
          percent=89.18,
          pct_lcb=87.28,
          pct_ucb=90.99) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Fair",
          study="NLA",
          percent=7.57,
          pct_lcb=6.03,
          pct_ucb=9.11) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Poor",
          study="NLA",
          percent=3.25,
          pct_lcb=2.21,
          pct_ucb=4.28) %>% 
  add_row(Indicator="Dissolved Oxygen", #NLA NAP prob
          Category="Good",
          study="NAP",
          percent=92.11,
          pct_lcb=87.00,
          pct_ucb=97.21) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Fair",
          study="NAP",
          percent=5.36,
          pct_lcb=0.85,
          pct_ucb=9.87) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Poor",
          study="NAP",
          percent=2.53,
          pct_lcb=0.08,
          pct_ucb=4.98) %>% 
  add_row(Indicator="Dissolved Oxygen", #ny prob
          Category="Good",
          study="NY",
          percent=91.39,
          pct_lcb=69.86,
          pct_ucb=112.92) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Fair",
          study="NY",
          percent=1.25,
          pct_lcb=0,
          pct_ucb=3.36) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Poor",
          study="NY",
          percent=7.36,
          pct_lcb=0,
          pct_ucb=18.59) %>% 
  add_row(Indicator="ChlA", #NLA results
          Category="Good",
          study="NLA",
          percent=35.22,
          pct_lcb=30.12,
          pct_ucb=40.33) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="NLA",
          percent=13.32,
          pct_lcb=10.11,
          pct_ucb=16.53) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="NLA",
          percent=51.45,
          pct_lcb=46.29,
          pct_ucb=56.62) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="NLA",
          percent=26.83,
          pct_lcb=22.07,
          pct_ucb=31.58) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="NLA",
          percent=21.53,
          pct_lcb=17.27,
          pct_ucb=25.79) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="NLA",
          percent=51.64,
          pct_lcb=46.40,
          pct_ucb=56.89) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="NLA",
          percent=26.96,
          pct_lcb=22.25,
          pct_ucb=31.68) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="NLA",
          percent=21.92,
          pct_lcb=17.38,
          pct_ucb=26.45) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="NLA",
          percent=51.12,
          pct_lcb=45.79,
          pct_ucb=56.45) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

forplot6$Category <- factor(forplot6$Category, levels=c("Good","Fair","Poor"))

ggplot(forplot6, aes(x=Category,fill=study)) +
  geom_point(aes(y=percent,color=study),stat = "identity", position=position_dodge(width=0.2))+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb,color=study), width=.25, position=position_dodge(width=0.2))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="EPA condition categories in  NLA, Northern Appalachian, NY, NH",y="Percent of Total",x="EPA Condition Classes")
```


## Sites map

```{r maps, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#read in lat/long for lakes
nh.locs <- att %>% 
  select(LAT_DD_N83,LON_DD_N83) %>% 
  rename(xcoord=LON_DD_N83,
         ycoord=LAT_DD_N83) %>% 
  mutate(study="NH")

ny.locs <- read.csv("Probability_Based_Sites_2020_2021.csv") %>% 
  select(LAT_DD83,LON_DD83) %>% 
  rename(xcoord=LON_DD83,
         ycoord=LAT_DD83) %>% 
  mutate(study="NY")

nap.locs <- read.csv("nla_2017_site_information-data.csv") %>% 
  filter(AREA_HA > 2.63,
         AG_ECO9 == "NAP") %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,LAT_DD83,LON_DD83) %>% 
  rename(ycoord=LAT_DD83,
         xcoord=LON_DD83) %>% 
  select(xcoord,ycoord) %>% 
  mutate(study="NAP")

nla.locs <- read.csv("nla_2017_site_information-data.csv") %>% 
  filter(AREA_HA > 2.63) %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,LAT_DD83,LON_DD83) %>% 
  rename(ycoord=LAT_DD83,
         xcoord=LON_DD83) %>% 
  select(xcoord,ycoord) %>% 
  mutate(study="NLA")

locs <- rbind(nh.locs,ny.locs,nla.locs,nap.locs)

#fetch map
library(ggmap)
library(maps)

states <- map_data("state")

locs.map <- ggplot() + 
  geom_polygon(data=states,aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3) +
  geom_point(data=locs,aes(x=xcoord,y=ycoord,color=study),shape=4, size=0.5)+
  guides(fill=FALSE)  # do this to leave off the color legend

print(locs.map)

NH <- map_data("state") %>% 
  filter(region=="new hampshire")

locs.map <- ggplot() + 
  geom_polygon(data=NH,aes(x = long, y = lat, group = group), fill = "white") + 
  coord_fixed(1.3) +
  geom_point(data=nh.locs,aes(x=xcoord,y=ycoord),shape=4, size=1)+
  guides(fill=FALSE)  # do this to leave off the color legend

print(locs.map)
```
