---
title: "Probability Sampling Results NLA 2017"
author: "Alene Onion and Tony Olsen of USEPA (modified by Sabrina Xie)"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y, %X')`"
output: html_document
---

```{r GlobalOptions}
options(knitr.duplicate.label = 'allow')
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# {.tabset}
Doing the probability sampling results analysis with NLA data of lakes in Northern Appalachians with area >6.5 acres

## Prepping data

Merging relevant raw data files and filtering

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
################################################################################################################
#reading in and formatting raw data
################################################################################################################

library(tidyverse)
library(huxtable)
library(ggplot2)
library(lubridate)

# read in NLA data
setwd("~/OneDrive - New York State Office of Information Technology Services/Rscripts/Probability.Sampling/")

chem <- read.csv("nla_2017_water_chemistry_chla-data.csv")

chem <- chem %>% 
  filter(ANALYTE %in% c("NTL","CHLA","PTL"),
         NARS_FLAG == "") %>% 
  select(UID,SITE_ID,DATE_COL,ANALYTE,RESULT) %>% 
  pivot_wider(names_from= ANALYTE ,values_from = RESULT,) %>% 
  rename(TN=NTL,TP=PTL)

oxygen <- read.csv("nla_2017_profile-data.csv")

oxygen <- oxygen %>% 
  select(UID,SITE_ID,DATE_COL,DEPTH,OXYGEN) %>% 
  # dissolved oxygen:
  # keep DO values between 0 and 2m depth
  mutate(keep=case_when(
    DEPTH <= 2 ~ "yes",
    DEPTH > 2 ~ "no")) %>%
  filter(keep=="yes") %>%
  select(-keep)

# mean DO for each lake
DO <- aggregate(oxygen$OXYGEN,list(oxygen$SITE_ID,oxygen$DATE_COL),FUN=mean)
# add back in
oxygen <- merge(oxygen,DO,by.x=c("SITE_ID","DATE_COL"),by.y=c("Group.1","Group.2"),all.x=TRUE)
oxygen$x <- as.numeric(oxygen$x)
oxygen <- oxygen %>% 
  select(-c(DEPTH,OXYGEN)) %>%
  rename(OXYGEN=x) %>% 
  distinct()

sites <- read.csv("nla_2017_site_information-data.csv")

sites <- sites %>% 
  mutate(include=case_when(
    AG_ECO9=="NAP" & AREA_HA>2.63 ~ "yes",
    SITE_ID=TRUE ~ "no"
  )) %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,YCOORD,XCOORD,WGT_TP_EXTENT,include) %>% 
  filter(WGT_TP_EXTENT != 0)

att <- merge(chem,oxygen,by=c("UID","SITE_ID","DATE_COL"),all.y=TRUE,all.x=TRUE)
att <- merge(att,sites,by=c("UID","SITE_ID"))
att$CHLA <- as.numeric(att$CHLA)
att$TN <- as.numeric(att$TN)
att$TP <- as.numeric(att$TP)
att$OXYGEN <- as.numeric(att$OXYGEN)
  
# selecting parameters and adding trophic status

att<-att %>% 
  # trophic status
  mutate(phos_trophic=case_when(
    TP<=10 ~ "oligotrophic",
    between(TP,10,20) ~ "mesotrophic",
    TP>=20 ~ "eutrophic")) %>%
  mutate(chla_trophic=case_when(
    CHLA<=2 ~ "oligotrophic",
    between(CHLA,2,8) ~ "mesotrophic",
    CHLA>=8 ~ "eutrophic")) %>%
  # EPA thresholds
  mutate(TP_threshold=case_when(
    TP<=16 ~ "Good",
    between(TP, 16, 27.9) ~ "Fair",
    TP>=27.9 ~ "Poor")) %>%
  mutate(TN_threshold=case_when(
    TN<=0.428 ~ "Good",
    between(TN, 0.428, 0.655) ~ "Fair",
    TN>=0.655 ~ "Poor")) %>%
  mutate(CHLA_threshold=case_when(
    CHLA<=4.52 ~ "Good",
    between(CHLA, 4.52, 8.43) ~ "Fair",
    CHLA>=8.43 ~ "Poor")) %>%
  # dissolved oxygen
  mutate(d.oxygen=case_when(
    OXYGEN<=3 ~ "Poor",
    between(OXYGEN, 3, 5) ~ "Fair",
    OXYGEN>=5 ~ "Good"))

framesize_LC <- c("EPA"=5721) 

```

## Comparison to NYS thresholds

The following code actually performs the analysis. It uses the number of sampled lakes in each trophic category (both for phosphorus and chlorophyll) to estimate the ratio of the 2583 lakes we could have sampled are eutrophic, mesotrophic, or eutrophic. 


PLEAE NOTE: this presumes that the lakes we didn't sample were random. A reasonable assumption but it's obviously preferable to sample all the lakes in the list to avoid this uncertainty. 

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
# Trophic classes
# Estimate assessment result

library(spsurvey)

CatExtent2 <- cat_analysis(
  dframe=att,
  vars=c("chla_trophic","phos_trophic"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

forplot2<-CatExtent2 %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Category!="Total",
         Subpopulation=="yes") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="chla_trophic","Trophic Status from ChlA",Indicator),
         Indicator=ifelse(Indicator=="phos_trophic","Trophic Status from Phosphorus",Indicator))

table <- forplot2 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

forplot2$Category <- factor(forplot2$Category, levels=c("oligotrophic", "mesotrophic", "eutrophic"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of trophic classes across Eastern Highland ponded waters",y="Percent of Total",x="Trophic status")

forplot5<-forplot2 %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NLA") %>% 
  add_row(Indicator="Trophic Status from ChlA", #lmas
          Category="eutrophic",
          study="LMAS",
          percent=((215/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="mesotrophic",
          study="LMAS",
          percent=((182/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="oligotrophic",
          study="LMAS",
          percent=((76/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="eutrophic",
          study="LMAS",
          percent=((211/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="mesotrophic",
          study="LMAS",
          percent=((141/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="oligotrophic",
          study="LMAS",
          percent=((121/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA", #ny prob
          Category="eutrophic",
          study="NYProb",
          percent=9.18,
          pct_lcb=2.88,
          pct_ucb=15.48) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="mesotrophic",
          study="NYProb",
          percent=70.32,
          pct_lcb=45.82,
          pct_ucb=94.82) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="oligotrophic",
          study="NYProb",
          percent=20.50,
          pct_lcb=5.86,
          pct_ucb=35.13) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="eutrophic",
          study="NYProb",
          percent=22.48,
          pct_lcb=9.56,
          pct_ucb=35.40)  %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="mesotrophic",
          study="NYProb",
          percent=32.08,
          pct_lcb=12.77,
          pct_ucb=51.38) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="oligotrophic",
          study="NYProb",
          percent=45.44,
          pct_lcb=27.37,
          pct_ucb=63.52) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

ggplot(forplot5, aes(x=Category,fill=study)) +
  geom_bar(aes(y=percent),stat = "identity", position = position_dodge())+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb), width=.05,
                 position=position_dodge(.9))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Trophic Classes in NLA vs. NY Probability vs. LMAS",y="Percent of Total",x="Trophic Classes")
  
```

## Comparison to NLA thresholds

The following code performs the analysis against EPA's condition thresholds as used in the National Lakes Assessments. It uses the number of sampled lakes in each condition category (for total phosphorus, total nitrogen and chlorophyll) to estimate the ratio of all lakes in good, fair and poor condition.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

# EPA nutrient threshold
#  Estimate assessment result
CatExtent <- cat_analysis(
  dframe=att,
  vars=c("TP_threshold","CHLA_threshold","TN_threshold"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

forplot<-CatExtent %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Category!="Total",Subpopulation=="yes") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="CHLA_threshold","ChlA",Indicator),
         Indicator=ifelse(Indicator=="TP_threshold","Phosphorus",Indicator),
         Indicator=ifelse(Indicator=="TN_threshold","Total Nitrogen",Indicator))

table <- forplot %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good", "Fair", "Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of Northern Appalachians ponded waters across EPA thresholds",y="Percent of Total",x="EPA threshold")
```

>>>>>>> 0ba55f3d292ab1d29f6a9760447658497f235422
Below, the analysis is done for each dissolved oxygen condition category.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
# Dissolved oxygen
#  Estimate assessment result
DOextent <- cat_analysis(
  dframe=att,
  vars=c("d.oxygen"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

#quickly calculate the percentages
totals<-DOextent %>% 
  select(Indicator,Category,Estimate.U) %>% 
  rename(total=Estimate.U)
totals$total <- DOextent$Estimate.U[4]
DOextent<-merge(DOextent,totals,by=c('Indicator','Category'),all.x = TRUE)

forplot4<-DOextent %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Subpopulation=="yes",Category!="Total") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="d.oxygen","Dissolved Oxygen",Indicator))

table <- forplot4 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good","Fair","Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of dissolved oxygen conditions across Northern Appalachians ponded waters",y="Percent of total",x="Condition category")
```

## Comparison to NY probability and LMAS

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#Plotting data frame distribution as well as NYS sample distribution collected since 2010
#LMAS numbers retrieved in 2020

frame<-att %>% 
  select(AREA_CAT6) %>% 
  rename(size_category=AREA_CAT6) %>% 
  distinct() %>%
  mutate(NLA_Sites = case_when(
    endsWith(size_category, "(1,4]") ~ ((40/490)*100),
    endsWith(size_category, "(4,10]") ~ ((96/490)*100),
    endsWith(size_category, "(10,20]") ~ ((115/490)*100),
    endsWith(size_category, "(20,50]") ~ ((105/490)*100),
    endsWith(size_category, ">50") ~ ((134/490)*100))
  ) %>%
  mutate(Probability_Sites = case_when(
    endsWith(size_category, "(1,4]") ~ ((1828/6437)*100),
    endsWith(size_category, "(4,10]") ~ ((2490/6437)*100),
    endsWith(size_category, "(10,20]") ~ ((1003/6437)*100),
    endsWith(size_category, "(20,50]") ~ ((616/6437)*100),
    endsWith(size_category, ">50") ~ ((500/6437)*100))
    ) %>%
  mutate(LMAS_Sites = case_when(
    endsWith(size_category, "(1,4]") ~ (7.64),
    endsWith(size_category, "(4,10]") ~ (16.7),
    endsWith(size_category, "(10,20]") ~ (16.2),
    endsWith(size_category, "(20,50]") ~ (19.3),
    endsWith(size_category, ">50") ~ (40.1)
  )) %>%
  ungroup() %>%
  gather(study,percentage,-size_category) %>% 
  arrange(study,size_category) %>% 
  mutate(percentage=as.numeric(percentage),
         size_category=factor(size_category,levels=c("(1,4]","(4,10]","(10,20]","(20,50]",">50")))

ggplot(frame, aes(x=study,y=percentage,fill=size_category)) +
  geom_bar(stat = "identity", position="stack")+
  ylim(0,100)+
  labs(title="Size classes in NLA vs. NY Probability vs. LMAS",y="Percent of Total",x="Size Category (hectares)")



```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

forplot6<-forplot %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NLA") %>% 
  add_row(Indicator="ChlA", #lmas results
          Category="Good",
          study="LMAS",
          percent=((306/752)*100)) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="LMAS",
          percent=((144/752)*100)) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="LMAS",
          percent=((302/752)*100)) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="LMAS",
          percent=((262/501)*100)) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="LMAS",
          percent=((82/501)*100)) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="LMAS",
          percent=((157/501)*100)) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="LMAS",
          percent=((186/441)*100)) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="LMAS",
          percent=((108/441)*100)) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="LMAS",
          percent=((147/441)*100)) %>% 
  add_row(Indicator="ChlA", #nys prob results
          Category="Good",
          study="NYProb",
          percent=50.01,
          pct_lcb=33.05,
          pct_ucb=66.97) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="NYProb",
          percent=40.81,
          pct_lcb=16.57,
          pct_ucb=65.05) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="NYProb",
          percent=9.18,
          pct_lcb=2.88,
          pct_ucb=15.48) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="NYProb",
          percent=65.59,
          pct_lcb=42.85,
          pct_ucb=88.33) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="NYProb",
          percent=15.44,
          pct_lcb=5.85,
          pct_ucb=25.04) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="NYProb",
          percent=18.97,
          pct_lcb=6.22,
          pct_ucb=31.72) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="NYProb",
          percent=61.49,
          pct_lcb=37.27,
          pct_ucb=85.71) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="NYProb",
          percent=14.95,
          pct_lcb=4.34,
          pct_ucb=25.56) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="NYProb",
          percent=23.57,
          pct_lcb=6.99,
          pct_ucb=40.14) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

forplot6$Category <- factor(forplot6$Category, levels=c("Good","Fair","Poor"))

ggplot(forplot6, aes(x=Category,fill=study)) +
  geom_bar(aes(y=percent),stat = "identity", position = position_dodge())+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb), width=.05,
                 position=position_dodge(.9))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="EPA condition categories in  NLA vs. NY Probability vs. LMAS",y="Percent of Total",x="EPA Condition Classes")
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
forplot7<-forplot4 %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NLA") %>% 
  add_row(Indicator="d.oxygen", #lmas
          Category="Good",
          study="LMAS",
          percent=((400/486)*100)) %>% 
  add_row(Indicator="d.oxygen",
          Category="Fair",
          study="LMAS",
          percent=((76/486)*100)) %>% 
  add_row(Indicator="d.oxygen",
          Category="Poor",
          study="LMAS",
          percent=((10/486)*100)) %>% 
  add_row(Indicator="d.oxygen", #ny prob
          Category="Good",
          study="NYProb",
          percent=46.89,
          pct_lcb=28.12,
          pct_ucb=65.65) %>% 
  add_row(Indicator="d.oxygen",
          Category="Fair",
          study="NYProb",
          percent=20.45,
          pct_lcb=7.86,
          pct_ucb=33.04) %>% 
  add_row(Indicator="d.oxygen",
          Category="Poor",
          study="NYProb",
          percent=32.66,
          pct_lcb=17.06,
          pct_ucb=48.26) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

forplot7$Category <- factor(forplot7$Category, levels=c("Good","Fair","Poor"))

ggplot(forplot7, aes(x=Category,fill=study)) +
  geom_bar(aes(y=percent),stat = "identity", position = position_dodge())+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb), width=.05,
                 position=position_dodge(.9))+
  ylim(0,100)+
  labs(title="Dissolved oxygen condition in  NLA vs. NY Probability vs. LMAS",y="Percent of Total",x="D.O. Condition")
```

## Sites map

```{r maps, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#read in lat/long for lakes
ny.locs <- read.csv("Probability_Based_Sites_2020_2021.csv") %>% 
  filter(EvalStatus=="Target_Sampled ") %>% 
  select(LAT_DD83,LON_DD83) %>% 
  rename(YCOORD=LAT_DD83,
         XCOORD=LON_DD83) %>% 
  mutate(study="NY")
nla.locs <- read.csv("nla_2017_site_information-data.csv") %>% 
  filter(AG_ECO9=="NAP",AREA_HA>2.63) %>% 
  select(LAT_DD83,LON_DD83) %>%
  rename(YCOORD=LAT_DD83,
         XCOORD=LON_DD83) %>% 
  mutate(study="NLA")

locs <- rbind(ny.locs,nla.locs)

#fetch map
library(ggmap)
library(maps)

states <- map_data("state")
east.coast <- subset(states, region %in% c("maine", "ohio", "new york", "pennsylvania","new jersey","connecticut","rhode island","massachusetts","vermont","new hampshire"))

locs.map <- ggplot() + 
  geom_polygon(data=east.coast,aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3) +
  geom_point(data=locs,aes(x=XCOORD,y=YCOORD,color=study))+
  guides(fill=FALSE)  # do this to leave off the color legend

print(locs.map)
```

## Questions

- Why are population sizes across size categories the same?

- Why are our error bars bigger than theirs?

- Should the weight adjustment be calculated using the whole population or the filtered subset?
