---
title: "Probability Sampling Results 2021"
author: "Alene Onion and Tony Olsen of USEPA (modified by Sabrina Xie)"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y, %X')`"
output: html_document
---

```{r GlobalOptions}
options(knitr.duplicate.label = 'allow')
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# {.tabset}
Doing the probability sampling results analysis with NLA data of lakes in Northern Appalachians with area >6.5 acres

## Prepping data

Merging relevant raw data files and filtering

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
################################################################################################################
#reading in and formatting raw data
################################################################################################################

library(tidyverse)
library(huxtable)
library(ggplot2)
library(lubridate)

# read in NLA data
setwd("~/OneDrive - New York State Office of Information Technology Services/Rscripts/Probability.Sampling/")

chem <- read.csv("nla_2017_water_chemistry_chla-data.csv")

chem <- chem %>% 
  filter(ANALYTE %in% c("NTL","CHLA","PTL","NITRATE_N","COLOR","CHLORIDE"),
         NARS_FLAG == "") %>% 
  select(UID,SITE_ID,DATE_COL,ANALYTE,RESULT) %>% 
  pivot_wider(names_from= ANALYTE ,values_from = RESULT,) %>% 
  rename(TN=NTL,TP=PTL)

oxygen <- read.csv("nla_2017_profile-data.csv")

oxygen <- oxygen %>% 
  select(UID,SITE_ID,DATE_COL,DEPTH,OXYGEN) %>% 
  # dissolved oxygen:
  # keep DO values between 0 and 2m depth
  mutate(keep=case_when(
    DEPTH <= 2 ~ "yes",
    DEPTH > 2 ~ "no")) %>%
  filter(keep=="yes") %>%
  select(-keep)

# mean DO for each lake
DO <- aggregate(oxygen$OXYGEN,list(oxygen$SITE_ID,oxygen$DATE_COL),FUN=mean)
# add back in
oxygen <- merge(oxygen,DO,by.x=c("SITE_ID","DATE_COL"),by.y=c("Group.1","Group.2"),all.x=TRUE)
oxygen$x <- as.numeric(oxygen$x)
oxygen <- oxygen %>% 
  select(-c(DEPTH,OXYGEN)) %>%
  rename(OXYGEN=x) %>% 
  distinct()

sites <- read.csv("nla_2017_site_information-data.csv")

sites <- sites %>% 
  mutate(include=case_when(
    AREA_HA>2.63 ~ "yes",
    SITE_ID=TRUE ~ "no"
  )) %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,YCOORD,XCOORD,WGT_TP_EXTENT,include) %>% 
  filter(WGT_TP_EXTENT != 0)

att <- merge(chem,oxygen,by=c("UID","SITE_ID","DATE_COL"),all.y=TRUE,all.x=TRUE)
att <- merge(att,sites,by=c("UID","SITE_ID"))
att$CHLA <- as.numeric(att$CHLA)
att$TN <- as.numeric(att$TN)
att$TP <- as.numeric(att$TP)
att$OXYGEN <- as.numeric(att$OXYGEN)
att$COLOR <- as.numeric(att$COLOR)
att$DINTP <- as.numeric(att$NITRATE_N)/(as.numeric(att$TP)/1000)
  
# selecting parameters and adding trophic status

att<-att %>% 
  # trophic status
  mutate(phos_trophic=case_when(
    TP<=10 ~ "oligotrophic",
    between(TP,10,20) ~ "mesotrophic",
    TP>=20 ~ "eutrophic")) %>%
  mutate(chla_trophic=case_when(
    CHLA<=2 ~ "oligotrophic",
    between(CHLA,2,8) ~ "mesotrophic",
    CHLA>=8 ~ "eutrophic")) %>%
  # EPA thresholds
  mutate(TP_threshold=case_when(
    TP<=16 ~ "Good",
    between(TP, 16, 27.9) ~ "Fair",
    TP>=27.9 ~ "Poor")) %>%
  mutate(TN_threshold=case_when(
    TN<=0.428 ~ "Good",
    between(TN, 0.428, 0.655) ~ "Fair",
    TN>=0.655 ~ "Poor")) %>%
  mutate(CHLA_threshold=case_when(
    CHLA<=4.52 ~ "Good",
    between(CHLA, 4.52, 8.43) ~ "Fair",
    CHLA>=8.43 ~ "Poor")) %>%
  # dissolved oxygen
  mutate(d.oxygen=case_when(
    OXYGEN<=3 ~ "Poor",
    between(OXYGEN, 3, 5) ~ "Fair",
    OXYGEN>=5 ~ "Good")) %>% 
  #leech
  mutate(leech=case_when(
    TP<=30 & COLOR<=20 ~ "Blue",
    TP>30 & COLOR<=20 ~ "Green",
    TP<=30 & COLOR>20 ~ "Brown",
    TP>30 & COLOR>20 ~ "Murky"
  )) %>% 
  #chloride
  mutate(chloride=case_when(
    CHLORIDE <= 35 ~ "Low",
    between(CHLORIDE,35,100) ~ "Medium",
    CHLORIDE >= 100 ~ "High"))

framesize_LC <- c("EPA"=4271) 

```

## Analysis

The start of the NYSAnalysis code from Tony Olsen

The results are weighted based on the distribution of the five size classes in the original data frame divided by the distribution in the samples evaluated. The purpose of this step is to account for oversampling in certain size classes.

## Comparison to NYS thresholds

The following code actually performs the analysis. It uses the number of sampled lakes in each trophic category (both for phosphorus and chlorophyll) to estimate the ratio of the 2583 lakes we could have sampled are eutrophic, mesotrophic, or eutrophic. 


PLEAE NOTE: this presumes that the lakes we didn't sample were random. A reasonable assumption but it's obviously preferable to sample all the lakes in the list to avoid this uncertainty. 

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

library(spsurvey)

# Size classes
SizeExtent <- cat_analysis(
  dframe=att,
  vars=c("AREA_CAT6"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")
table <- SizeExtent %>% 
  filter(Subpopulation=="yes") %>%
  select(Category,nResp,Estimate.U,LCB95Pct.U,UCB95Pct.U)
hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

# Trophic classes
# Estimate assessment result
CatExtent2 <- cat_analysis(
  dframe=att,
  vars=c("chla_trophic","phos_trophic"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

forplot2<-CatExtent2 %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Subpopulation=="yes",Category!="Total") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="chla_trophic","Trophic Status from ChlA",Indicator),
         Indicator=ifelse(Indicator=="phos_trophic","Trophic Status from Phosphorus",Indicator))

table <- forplot2 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

forplot2$Category <- factor(forplot2$Category, levels=c("oligotrophic", "mesotrophic", "eutrophic"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of trophic classes across Eastern Highland ponded waters",y="Percent of Total",x="Trophic status")

forplot5<-forplot2 %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NLA") %>% 
  add_row(Indicator="Trophic Status from ChlA", #lmas
          Category="eutrophic",
          study="LMAS",
          percent=((215/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="mesotrophic",
          study="LMAS",
          percent=((182/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="oligotrophic",
          study="LMAS",
          percent=((76/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="eutrophic",
          study="LMAS",
          percent=((211/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="mesotrophic",
          study="LMAS",
          percent=((141/473)*100)) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="oligotrophic",
          study="LMAS",
          percent=((121/473)*100)) %>% 
  add_row(Indicator="Trophic Status from ChlA", #ny prob
          Category="eutrophic",
          study="NYProb",
          percent=9.18,
          pct_lcb=2.88,
          pct_ucb=15.48) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="mesotrophic",
          study="NYProb",
          percent=70.32,
          pct_lcb=45.82,
          pct_ucb=94.82) %>% 
  add_row(Indicator="Trophic Status from ChlA",
          Category="oligotrophic",
          study="NYProb",
          percent=20.50,
          pct_lcb=5.86,
          pct_ucb=35.13) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="eutrophic",
          study="NYProb",
          percent=22.48,
          pct_lcb=9.56,
          pct_ucb=35.40)  %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="mesotrophic",
          study="NYProb",
          percent=32.08,
          pct_lcb=12.77,
          pct_ucb=51.38) %>% 
  add_row(Indicator="Trophic Status from Phosphorus",
          Category="oligotrophic",
          study="NYProb",
          percent=45.44,
          pct_lcb=27.37,
          pct_ucb=63.52) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

ggplot(forplot5, aes(x=Category)) +
  geom_point(aes(y=percent,color=study),stat = "identity", position=position_dodge(width=0.2))+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb,color=study), width=.25, position=position_dodge(width=0.2))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Trophic Classes in NLA vs. NY Probability vs. LMAS",y="Percent of Total",x="Trophic Classes")

##### CHLORIDE #####
CatExtent <- cat_analysis(
  dframe=att,
  vars=c("chloride"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

forplot.chloride<-CatExtent %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Category!="Total",Subpopulation=="yes") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category)

table <- forplot.chloride %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)


```

## Comparison to NLA thresholds

The following code performs the analysis against EPA's condition thresholds as used in the National Lakes Assessments. It uses the number of sampled lakes in each condition category (for total phosphorus, total nitrogen and chlorophyll) to estimate the ratio of all lakes in good, fair and poor condition.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}

# EPA nutrient threshold
#  Estimate assessment result
CatExtent <- cat_analysis(
  dframe=att,
  vars=c("TP_threshold","CHLA_threshold","TN_threshold"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

forplot<-CatExtent %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Subpopulation=="yes",Category!="Total") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="CHLA_threshold","ChlA",Indicator),
         Indicator=ifelse(Indicator=="TP_threshold","Phosphorus",Indicator),
         Indicator=ifelse(Indicator=="TN_threshold","Total Nitrogen",Indicator))

table <- forplot %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good", "Fair", "Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of Northern Appalachians ponded waters across EPA thresholds",y="Percent of Total",x="EPA threshold")
```

Below, the analysis is done for each dissolved oxygen condition category.

```{r, echo=TRUE, results='asis', message=FALSE, warning=FALSE}
# Dissolved oxygen
#  Estimate assessment result
DOextent <- cat_analysis(
  dframe=att,
  vars=c("d.oxygen"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

#quickly calculate the percentages
totals<-DOextent %>% 
  select(Indicator,Category,Estimate.U) %>% 
  rename(total=Estimate.U)
totals$total <- DOextent$Estimate.U[4]
DOextent<-merge(DOextent,totals,by=c('Indicator','Category'),all.x = TRUE)

forplot4<-DOextent %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>%
  filter(Category!="Total",Subpopulation=="yes") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="d.oxygen","Dissolved Oxygen",Indicator))

table <- forplot4 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

table$Category <- factor(table$Category, levels=c("Good","Fair","Poor"))

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Distribution of dissolved oxygen conditions across Northern Appalachians ponded waters",y="Percent of total",x="Condition category")
```

## Comparison to NY probability and LMAS

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#Plotting data frame distribution as well as NYS sample distribution collected since 2010
#LMAS numbers retrieved in 2020

frame<-att %>% 
  select(AREA_CAT6) %>% 
  rename(size_category=AREA_CAT6) %>% 
  distinct() %>%
  mutate(NAP = case_when(
    endsWith(size_category, "(1,4]") ~ ((1870/9727)*100),
    endsWith(size_category, "(4,10]") ~ ((2380/9727)*100),
    endsWith(size_category, "(10,20]") ~ ((1791/9727)*100),
    endsWith(size_category, "(20,50]") ~ ((1695/9727)*100),
    endsWith(size_category, ">50") ~ ((1991/9727)*100))
  ) %>%
  mutate(NY = case_when(
    endsWith(size_category, "(1,4]") ~ ((1828/6437)*100),
    endsWith(size_category, "(4,10]") ~ ((2490/6437)*100),
    endsWith(size_category, "(10,20]") ~ ((1003/6437)*100),
    endsWith(size_category, "(20,50]") ~ ((616/6437)*100),
    endsWith(size_category, ">50") ~ ((500/6437)*100))
    ) %>%
  mutate(NLA = case_when(
    endsWith(size_category, "(1,4]") ~ ((15723.70/72957.58)*100),
    endsWith(size_category, "(4,10]") ~ ((20956.61/72957.58)*100),
    endsWith(size_category, "(10,20]") ~ ((13494.61/72957.58)*100),
    endsWith(size_category, "(20,50]") ~ ((11130.79/72957.58)*100),
    endsWith(size_category, ">50") ~ ((11651.86/72957.58)*100))) %>%
  ungroup() %>%
  gather(study,percentage,-size_category) %>% 
  arrange(study,size_category) %>% 
  mutate(percentage=as.numeric(percentage),
         size_category=factor(size_category,levels=c("(1,4]","(4,10]","(10,20]","(20,50]",">50")))

ggplot(frame, aes(x=study,y=percentage,fill=size_category)) +
  geom_bar(stat = "identity", position="stack")+
  ylim(0,100)+
  labs(title="Size classes in NLA national vs. Northern Appalachian vs. NY",y="Percent of Total",x="Size Category (hectares)")


```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

forplot6<-forplot %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NLA") %>% 
  add_row(Indicator="ChlA", #NAP results
          Category="Good",
          study="NAP",
          percent=73.22,
          pct_lcb=64.45,
          pct_ucb=81.99) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="NAP",
          percent=7.96,
          pct_lcb=3.68,
          pct_ucb=12.23) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="NAP",
          percent=18.82,
          pct_lcb=11.26,
          pct_ucb=26.38) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="NAP",
          percent=64.20,
          pct_lcb=53.89,
          pct_ucb=74.51) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="NAP",
          percent=20.81,
          pct_lcb=12.76,
          pct_ucb=28.85) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="NAP",
          percent=15.00,
          pct_lcb=8.25,
          pct_ucb=21.74) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="NAP",
          percent=66.41,
          pct_lcb=55.32,
          pct_ucb=77.50) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="NAP",
          percent=25.15,
          pct_lcb=13.57,
          pct_ucb=36.73) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="NAP",
          percent=8.44,
          pct_lcb=4.23,
          pct_ucb=12.65) %>% 
  add_row(Indicator="ChlA", #nys prob results
          Category="Good",
          study="NY",
          percent=50.01,
          pct_lcb=33.05,
          pct_ucb=66.97) %>% 
  add_row(Indicator="ChlA",
          Category="Fair",
          study="NY",
          percent=40.81,
          pct_lcb=16.57,
          pct_ucb=65.05) %>% 
  add_row(Indicator="ChlA",
          Category="Poor",
          study="NY",
          percent=9.18,
          pct_lcb=2.88,
          pct_ucb=15.48) %>% 
  add_row(Indicator="Phosphorus",
          Category="Good",
          study="NY",
          percent=65.59,
          pct_lcb=42.85,
          pct_ucb=88.33) %>% 
  add_row(Indicator="Phosphorus",
          Category="Fair",
          study="NY",
          percent=15.44,
          pct_lcb=5.85,
          pct_ucb=25.04) %>% 
  add_row(Indicator="Phosphorus",
          Category="Poor",
          study="NY",
          percent=18.97,
          pct_lcb=6.22,
          pct_ucb=31.72) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Good",
          study="NY",
          percent=61.49,
          pct_lcb=37.27,
          pct_ucb=85.71) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Fair",
          study="NY",
          percent=14.95,
          pct_lcb=4.34,
          pct_ucb=25.56) %>% 
  add_row(Indicator="Total Nitrogen",
          Category="Poor",
          study="NY",
          percent=23.57,
          pct_lcb=6.99,
          pct_ucb=40.14)  %>% 
  add_row(Indicator="Dissolved Oxygen", #NLA
          Category="Good",
          study="NLA",
          percent=89.18,
          pct_lcb=87.28,
          pct_ucb=90.99) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Fair",
          study="NLA",
          percent=7.57,
          pct_lcb=6.03,
          pct_ucb=9.11) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Poor",
          study="NLA",
          percent=3.25,
          pct_lcb=2.21,
          pct_ucb=4.28) %>% 
  add_row(Indicator="Dissolved Oxygen", #NLA NAP prob
          Category="Good",
          study="NAP",
          percent=92.11,
          pct_lcb=87.00,
          pct_ucb=97.21) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Fair",
          study="NAP",
          percent=5.36,
          pct_lcb=0.85,
          pct_ucb=9.87) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Poor",
          study="NAP",
          percent=2.53,
          pct_lcb=0.08,
          pct_ucb=4.98) %>% 
  add_row(Indicator="Dissolved Oxygen", #ny prob
          Category="Good",
          study="NY",
          percent=91.39,
          pct_lcb=69.86,
          pct_ucb=100) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Fair",
          study="NY",
          percent=1.25,
          pct_lcb=0,
          pct_ucb=3.36) %>% 
  add_row(Indicator="Dissolved Oxygen",
          Category="Poor",
          study="NY",
          percent=7.36,
          pct_lcb=0,
          pct_ucb=18.59) %>% 
  add_row(Indicator="chloride",
          Category="High",
          study="NAP",
          percent=23.42,
          pct_lcb=15.19,
          pct_ucb=31.64) %>% 
  add_row(Indicator="chloride",
          Category="Low",
          study="NAP",
          percent=60.34,
          pct_lcb=50.78,
          pct_ucb=69.90) %>% 
  add_row(Indicator="chloride",
          Category="Medium",
          study="NAP",
          percent=16.24,
          pct_lcb=9.67,
          pct_ucb=22.81) %>% 
  add_row(Indicator="chloride",
          Category="High",
          study="NY",
          percent=1.20,
          pct_lcb=0.00,
          pct_ucb=3.20) %>% 
  add_row(Indicator="chloride",
          Category="Low",
          study="NY",
          percent=82.94,
          pct_lcb=60.41,
          pct_ucb=100) %>%
  add_row(Indicator="chloride",
          Category="Medium",
          study="NY",
          percent=15.87,
          pct_lcb=4.73,
          pct_ucb=27.00) %>% 
  add_row(Indicator="chloride",
          Category="High",
          study="NLA",
          percent=25.41,
          pct_lcb=21.17,
          pct_ucb=29.65) %>%
  add_row(Indicator="chloride",
          Category="Low",
          study="NLA",
          percent=59.49,
          pct_lcb=54.44,
          pct_ucb=64.53) %>% 
  add_row(Indicator="chloride",
          Category="Medium",
          study="NLA",
          percent=15.10,
          pct_lcb=11.77,
          pct_ucb=18.43) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  
forplot6$Category <- factor(forplot6$Category, levels=c(c("Good","Fair","Poor"),c("Low", "Medium", "High")))

ggplot(forplot6, aes(x=Category,fill=study)) +
  geom_point(aes(y=percent,color=study),stat = "identity", position=position_dodge(width=0.2))+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb,color=study), width=.25, position=position_dodge(width=0.2))+
  facet_wrap(~Indicator,scales = "free")+
  ylim(0,100)+
  labs(title="Condition estimates in  NLA vs. Northern Appalachian vs. NY",y="Percent of Total",x="Condition Classes")
```

## Leech nutrient-color

```{r,echo=FALSE,results="asis",message=FALSE,warning=FALSE}

print(
  ggplot(att,aes(x=TP,y=COLOR,color=leech))+
    geom_point()+
    geom_vline(xintercept=30)+
    geom_hline(yintercept=20)+
    scale_y_continuous(trans='log10')+
    scale_x_continuous(trans='log10')+
    scale_color_manual(values=c("blue", "brown", "green","#E69F00")))

CatExtent2 <- cat_analysis(
  dframe=att,
  vars=c("leech"),
  subpops = "include",
  siteID = "SITE_ID",
  weight = "WGT_TP_EXTENT",
  xcoord = "XCOORD",
  ycoord = "YCOORD")

forplot2<-CatExtent2 %>% 
  mutate(pct=Estimate.P,
         pct_lcb=LCB95Pct.P,
         pct_ucb=UCB95Pct.P) %>% 
  filter(Category!="Total",
         Subpopulation=="yes") %>% 
  select(Indicator,Category,pct,pct_lcb,pct_ucb) %>% 
  distinct() %>% 
  gather(percent,results,-Indicator,-Category) %>% 
  mutate(Indicator=ifelse(Indicator=="leech","Lake nutrient-color status",Indicator))

table <- forplot2 %>% 
  spread(percent,results)  %>% 
  rename(Percentage=pct,Lower95Confidence=pct_lcb,Upper95Confidence=pct_ucb)

hux <- as_hux(table)
number_format(hux) <- 2
theme_plain(hux)

ggplot(table,aes(x=Category,y=Percentage)) +
  geom_point()+
  geom_errorbar(aes(ymin=Lower95Confidence,ymax=Upper95Confidence),width=0.2)+
  theme(legend.position = "none")+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Lake nutrient-color status across Northern Appalachian ponded waters",y="Percent of Total",x="Lake nutrient-color status")

forplot<-forplot2 %>% 
  spread(percent,results) %>% 
  rename(percent=pct) %>% 
  mutate(study="NLA") %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Blue",
          study="NAP",
          percent=45.15,
          pct_lcb=33.41,
          pct_ucb=56.90) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Brown",
          study="NAP",
          percent=46.93,
          pct_lcb=35.32,
          pct_ucb=58.54) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Green",
          study="NAP",
          percent=1.95,
          pct_lcb=0.76,
          pct_ucb=3.15) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Murky",
          study="NAP",
          percent=5.96,
          pct_lcb=2.37,
          pct_ucb=9.56) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Blue",
          study="NYS",
          percent=20.73,
          pct_lcb=6.73,
          pct_ucb=34.74) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Brown",
          study="NYS",
          percent=52.22,
          pct_lcb=20.26,
          pct_ucb=84.18) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Green",
          study="NYS",
          percent=1.22,
          pct_lcb=0.00,
          pct_ucb=3.35) %>% 
  add_row(Indicator="Lake nutrient-color status",
          Category="Murky",
          study="NYS",
          percent=25.83,
          pct_lcb=5.51,
          pct_ucb=46.16) %>% 
  mutate(pct_lcb=ifelse(is.na(pct_lcb),percent,pct_lcb),
         pct_ucb=ifelse(is.na(pct_ucb),percent,pct_ucb))
  

ggplot(forplot, aes(x=Category,fill=study)) +
  geom_point(aes(y=percent,color=study),stat = "identity", position=position_dodge(width=0.2))+
  geom_errorbar(aes(ymin=pct_lcb, ymax=pct_ucb,color=study), width=.25, position=position_dodge(width=0.2))+
  facet_wrap(~Indicator,scales = "free_y")+
  ylim(0,100)+
  labs(title="Lake nutrient-color status in  NLA vs. Northern Appalachian vs. NY",y="Percent of Total",x="Lake nutrient-color status")

```

## Sites map

```{r maps, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#read in lat/long for lakes
ny.locs <- read.csv("Probability_Based_Sites_2020_2021.csv") %>% 
  select(LAT_DD83,LON_DD83) %>% 
  rename(xcoord=LON_DD83,
         ycoord=LAT_DD83) %>% 
  mutate(study="NY")

nap.locs <- read.csv("nla_2017_site_information-data.csv") %>% 
  filter(AREA_HA > 2.63,
         AG_ECO9 == "NAP") %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,LAT_DD83,LON_DD83) %>% 
  rename(ycoord=LAT_DD83,
         xcoord=LON_DD83) %>% 
  select(xcoord,ycoord) %>% 
  mutate(study="NAP")

nla.locs <- read.csv("nla_2017_site_information-data.csv") %>% 
  filter(AREA_HA > 2.63) %>% 
  select(UID,SITE_ID,AREA_CAT6,EVAL_CAT,LAT_DD83,LON_DD83) %>% 
  rename(ycoord=LAT_DD83,
         xcoord=LON_DD83) %>% 
  select(xcoord,ycoord) %>% 
  mutate(study="NLA")

locs <- rbind(ny.locs,nla.locs,nap.locs)

#fetch map
library(ggmap)
library(maps)

states <- map_data("state")

locs.map <- ggplot() + 
  geom_polygon(data=states,aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3) +
  geom_point(data=locs,aes(x=xcoord,y=ycoord,color=study),shape=4, size=0.5)+
  guides(fill=FALSE)  # do this to leave off the color legend

print(locs.map)
```
