---
title: "Probability Sampling Results 2021"
author: "Alene Onion and Tony Olsen of USEPA (updated by Sabrina Xie)"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y, %X')`"
output: html_document
---

```{r GlobalOptions}
options(knitr.duplicate.label = 'allow')
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(huxtable)
library(ggplot2)
library(lubridate)
```


This script uses EPA's SPsurvey package to estimate how many of NYS ponded waters have "excursions" from NY StayCALM water quality standards based on the results of 2020-2021 probability-based sampling.

```{r}

#read in data on excursions
excursions <- read.csv("probability.data.excursions.csv")
excursions <- excursions %>% 
  select(parameter,n_exceedances,LAKE_ID) %>% 
  spread(parameter,n_exceedances,fill = NA)

#read in data on sites
sites<-read.csv("Probability_Based_Sites_2020_2021.csv")
sites<-sites %>% 
  rename(siteID=SITE_ID,
         xcoord=LON_DD83,
         ycoord=LAT_DD83,
         Eval_Status=EvalStatus) %>% 
  #removing sites that we haven't yet evaluated
  filter(Eval_Status!="") %>% 
  mutate(Eval_Status=trimws(Eval_Status))

att <- merge(sites,excursions,by=c(SITE_ID=LAKE_ID))